---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

```{r load-data}
load("/Users/iBowen/Desktop/stat/stat_coursera/project2/gss.Rdata")
```


* * *

## Part 1: Data

Describe how the observations in the sample are collected, and the implications of this data collection method on the scope of inference (generalizability / causality).

As mentioned in the Codebook, the observation data is collected by computer-assisted personal interview (CAPI), face-to-face interview, and telephone interview. There is no experiment assignment occuring int sampling. Hence, the research question below could only produce generalizability but not causality.


* * *

## Part 2: Research question

```{r load-data}
str(gss)
```

* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

* * *

## Part 4: Inference

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.