---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

```{r load-data}
load("/Users/iBowen/Desktop/stat/stat_coursera/project2/gss.Rdata")
```


* * *

## Part 1: Data

__As mentioned in the Codebook, the observation data is collected by computer-assisted personal interview (CAPI), face-to-face interview, and telephone interview. There is no experiment assignment occuring int sampling. Hence, the research question below could only produce generalizability but not causality.__


* * *

## Part 2: Research question

__Does there appear to be a relationship between political party affiliation and
region of interview?__

* * *

## Part 3: Exploratory data analysis

```{r}
# subset political party affiliation as 'partyid' and region of interview as 'region'
data <- gss[, c('partyid', 'region')]
table(data)
# each cell has at least 5 cases

# make a mosaic plot to show the distribution under party and affiliation
mosaicplot(~ region + partyid, data = data, main = 'Distribution of Political Party Affiliation and Region of Interview', color = TRUE)

# note: I couldn't find a solution to handle label overlapping, but it won't affect to find the proportion difference inbetween each cell.
```

As shown in the plotting, the proportions are different along each variable. But before we confirm a dependent relationship between these two variables, an inference test is needed. Since we have two categorical varaibles, so a Chi-Square statistics is used to test the independence between party affiliation and region of interview.

* * *

## Part 4: Inference

__Step 1. Check the condition__

Independence:

- Random sample/assignment: the observations are collected by independent interview.
- If sampling without replacement, n < 10% of population. The total data points are 57,061, absolutly less than the 10% of US popluation.
- Each case only contributes to one cell in the table.

Sample size: 

- Each particular scenario must have at least 5 expected cases. The minimum cell is 41

So, it meets the condiiton for Chi-Square test

__Step 2. Build the Null and Alternative Hypothesis__

H0: Political affiliation and region of interview are independent. Political affiliation does not vary by region of interview.

HA: Political affiliation and region of interview are dependent. Political affiliation does vary by region of interview.

__Step 3. Implement the Chi-Square Independent Test__

```{r}
# transform table into dataframe with NA removed
matrix.obs <- as.data.frame.matrix(table(data))
# calculate rowsum and colsum
rowtotal <- rowSums(matrix.obs)
coltotal <- colSums(matrix.obs)
# record the total value
total <- sum(rowtotal)
# calculate expected matrix
matrix.exp <- (rowtotal / total) %*% t(coltotal)
# calculate X2 and degree of freedom
X2 <- sum((matrix.obs - matrix.exp) ^ 2 / matrix.exp)
df <- (dim(matrix.obs)[1] - 1) * (dim(matrix.obs)[2] - 1)
# calculate the p-value
pchisq(X2, df, lower.tail = FALSE)
```

__Conclusion__
we can find the p-value is extremely small compared the 5% significant level, so we are going to reject the H0, and accect the HA. That means, there is a associaiton between the region of interview and political affiliation.

